---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# `gecevar` R Package 


`gecevar` provides a set of climatic and environmental data for a choosen area/country that can be used for ecological analyses. Data come from a variety of sources Chelsa, SRTM, BRGM, and various research projects for the environment). Climatic and environmental data are available as multiband raster files at a choosen resolution (a least 250m) in the geographic projection you need.

## System requirements

Make sure Curl, GDAL and GRASS GIS are installed on your system.

## Installation

You can install **gecevar** from [GitHub](https://github.com/ghislainv/gecevar) with:

```{r github, eval=FALSE}
devtools::install_github("ghislainv/gecevar")
```

## Contributing

The `gecevar` R package is Open Source and released under the [GNU GPL version 3](https://www.gnu.org/licenses/gpl-3.0.en.html) license. Anybody who is interested can contribute to the package development following our [Community guidelines](articles/Contributing.html). Every contributor must agree to follow the project's [Code of Conduct](articles/Code_of_conduct.html).
